Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2016-01-17 cordillera-cycle.tex
Typeset: remove no-break space from abstract.
It is apparently not allowed.

\begin{abstract}
Despite more than a century of geological observations, the
Cordilleran ice sheet of North America remains poorly understood in
terms of its former extent, volume and dynamics. Although
geomorphological evidence is abundant, its complexity is such that
whole ice-sheet reconstructions of advance and retreat patterns are
lacking. Here we use a numerical ice sheet model calibrated against
field-based evidence to attempt a quantitative reconstruction of the
Cordilleran ice sheet history through the last glacial
cycle. A series of simulations is driven by time-dependent
temperature offsets from six proxy records located around the
globe. Although this approach reveals large variations in model
response to evolving climate forcing, all simulations produce two
major glaciations during marine oxygen isotope stages~4
(61.9--56.5\,\unit{ka}) and~2 (23.2--16.8\,\unit{ka}). The timing of
glaciation is better reproduced using temperature reconstructions
from Greenland and Antarctic ice cores than from regional oceanic
sediment cores. During most of the last glacial cycle, the modelled
ice cover is discontinuous and restricted to high mountain
areas. However, widespread precipitation over the Skeena Mountains
favours the persistence of a central ice dome throughout the glacial
cycle. It acts as a nucleation centre before the Last Glacial
Maximum and hosts the last remains of Cordilleran ice until the
qmiddle Holocene (6.6--6.2\,\unit{ka}).
\end{abstract}

>>>>

\begin{abstract}
Despite more than a century of geological observations, the
Cordilleran ice sheet of North America remains poorly understood in
terms of its former extent, volume and dynamics. Although
geomorphological evidence is abundant, its complexity is such that
whole ice-sheet reconstructions of advance and retreat patterns are
lacking. Here we use a numerical ice sheet model calibrated against
field-based evidence to attempt a quantitative reconstruction of the
Cordilleran ice sheet history through the last glacial
cycle. A series of simulations is driven by time-dependent
temperature offsets from six proxy records located around the
globe. Although this approach reveals large variations in model
response to evolving climate forcing, all simulations produce two
major glaciations during marine oxygen isotope stages 4
(61.9--56.5\,\unit{ka}) and 2 (23.2--16.8\,\unit{ka}). The timing of
glaciation is better reproduced using temperature reconstructions
from Greenland and Antarctic ice cores than from regional oceanic
sediment cores. During most of the last glacial cycle, the modelled
ice cover is discontinuous and restricted to high mountain
areas. However, widespread precipitation over the Skeena Mountains
favours the persistence of a central ice dome throughout the glacial
cycle. It acts as a nucleation centre before the Last Glacial
Maximum and hosts the last remains of Cordilleran ice until the
middle Holocene (6.6--6.2\,\unit{ka}).
\end{abstract}
